Computable Compressed Matrices\\Suplementary Material

### 2013-09-27 manuscript.tex
Tudo analisado, menos a conclusão

 Where \textit{bits alocated} above mean total bits required for standard storage of the matrix, without compression, while bits used mean total bits requires to store the matrix after compression. From now on the efficiencies are denoted by $\eta_1$ for the SM method and by $\eta_2$ for the VLB method. 
 \subsubsection*{SM Method}
 Let $M_{r \times c}$ be the matrix we wish to compress. In comparison with a conventional allocation (64-bit integers), we can apply Equation \ref{eq:14} to caluculate the efficiency of the SM method:

@@

As we see in \ref{eq:15}, $\eta_1$ does not depend on size of the matrix, only on the bit-length of $\max M$. If $b(\max M)=64$, $\eta_1$ is 0, i.e., no compression is possible. On the other extreme, if the matrix is composed exclusively of 0s and 1s, maximal compression is achievable, $\eta_1=1$.

@@

Let the $rc$ elements of the matrix $M_{r \times c}$ be divided into $g$ groups, each with $f_i$ numbers of bit-length $b_i = b(m_i)$. Thus $f_i$ is the frequency of each bit-length present in $M$. Let $k = b(b(\max M))$, i.e., the bit length of the bit-length of $\max M$. The efficiency $\eta_2$ is shown below.

@@

@@

\begin{equation}\label{eq:18}
 F(x) =  w \, Beta(\alpha_1,\beta_1) + (1-w)\,Beta(\alpha_2,\beta_2)
\end{equation}

The intention of using this mixture was to find a simple way to represent a large variety of bit length distributions. The first two central moments of this mixture are given in \ref{eq:19} and will be used later to summarize our numerical results.

@@

From now on we will apply Equations \ref{eq:15} and \ref{eq:17}, to determine the compression efficiency of SM and VLB methods for random matrices generated as describe above.

With $w=0$, a single Beta distribution is used. In Figure \ref{fig:01020304}, we show some distributions of bit-lengths for some combinations of $\alpha_1$ and $\beta_1$. From the figure it can be seen that a large variety of unimodal distributions can be generated in the interval $[1,64]$.

\begin{figure}[ht]
  \centering
  \subfigure[$\alpha=1,\beta_1=1$]{
%   \includegraphics[scale=0.25,angle=-90,clip]{fig01}
  }
  \subfigure[$\alpha=1,\beta=32$]{
%   \includegraphics[scale=0.25,angle=-90,clip]{fig02}
  }
  \subfigure[$\alpha=32,\beta=1$]{
%   \includegraphics[scale=0.25,angle=-90,clip]{fig03}
  }
  \subfigure[$\alpha=64,\beta=64$]{
%   \includegraphics[scale=0.25,angle=-90]{fig04}
  }
  \caption{Histograms constructed from samples with 10,000 elements, generated from a Beta distribution. Below each histogram is possible to verify the parameters used.}
  \label{fig:01020304}
\end{figure}

@@

\begin{figure}[h]
  \centering
   \subfigure[SM-VLB  Difference ($D = \eta_1 - \eta_2$)]{ 
%    \includegraphics[scale=0.18,clip]{fig07}
   }
   \subfigure[SM Efficiency  ($\eta_1$)]{
%    \includegraphics[scale=0.18,clip]{fig06}
   }
   \subfigure[VBL Efficiency ($\eta_2$)]{
%    \includegraphics[scale=0.18,clip]{fig05}
   }
   \subfigure[SM$\geq$VLB Efficiency ($\eta_1 \geq \eta_2$)]{
%    \includegraphics[scale=0.18,clip]{fig08}
   }
  \caption{Comparing compression efficiency of methods 1 and 2. Color scale in (a), (b) and (c) represent average bit-length. In (a) we can see the difference $D = \eta_1 - \eta_2$. It can be seen that for most combination of $\alpha$ and $\beta$, $D<0$, meaning the second method is more efficent to compress a sample of numbers with bit-lengths coming from a $Beta(\alpha,\beta)$ distribution. However, there is a small region in parameter space, which is shown in white on (d), where the SM method is more efficient. This region corresponds to the dots in red in (a), where the average bit-length is higher. In panels (b) and (c), we can see the efficiencies of SM and VLB methods, respectively.}
  \label{fig:06070809}
\end{figure}

Now let $w = 0.5$, i.e., matrices will have elements with bit-lengths comming from a mixture of beta distributions, $B_1\sim Beta(\alpha_1,\beta_1)$ and $B_2\sim Beta(\alpha_2,\beta_2)$. The expected value for this mixture is shown in Equation \ref{eq:22}.

@@

As before, we generate 3 matrices of dimension $10,000$ for each parameterization, calculate the average efficiencies (Equations \ref{eq:23} and \ref{eq:24}) and their diference $D$.

Before moving on to efficiency results and analyses, let's first inspect samples from the mixture of transformed Beta distributions. Figures \ref{fig:09101112} and \ref{fig:13141516}, show a few parameterizations and their resulting sample distributions. It is important to note that from the mixture we can now generate bimodal distributions as well as the unimodal types tested before. Since we are making statements about efficiency as a function of the expected bit-length, it is important to verify if these statements hold for bimodal  distributions as well.

After sampling uniformly ($[1,5,9,\ldots,64]$, $n=65,536$) the bit-length space and comparing efficiencies, we summarized the results on Table \ref{tab:02}. In it we see how many parameterizations (from our sample) favor each method. We can also look at the distribution of efficiencies on our samples for each method (Figure \ref{fig:1718}), which clearly demonstrate the greater expected efficiency of method VLB (Figure \ref{fig:18}).

\begin{figure}[h]
  \centering
  \subfigure[$\alpha_1=1,\beta_1=1,\alpha_2=1,\beta_2=1$]{
%   \includegraphics[scale=0.28,angle=-90,clip]{fig09}
  }
  \subfigure[$\alpha_1=1,\beta_1=32,\alpha_2=32,\beta_2=1$]{
%   \includegraphics[scale=0.28,angle=-90,clip]{fig10}
  }
  \subfigure[$\alpha_1=32,\beta_1=32,\alpha_2=32,\beta_2=32$]{
%   \includegraphics[scale=0.28,angle=-90,clip]{fig11}
  }
  \subfigure[$\alpha_1=64,\beta_1=32,\alpha_2=32,\beta_2=64$]{
%   \includegraphics[scale=0.28,angle=-90]{fig12}
  }
  \caption{Histograms constructed from samples with 10,000 elements, generated from the mixture of two Beta distributions with $w=0.5$. Below each histogram are the parameters of the mixture.}
  \label{fig:09101112}
\end{figure}

\begin{figure}[h]
  \centering
  \subfigure[$\alpha_1=64,\beta_1=48,\alpha_2=1,\beta_2=48$]{
%   \includegraphics[scale=0.28,angle=-90]{fig13}
  }
  \subfigure[$\alpha_1=16,\beta_1=46,\alpha_2=49,\beta_2=64$]{
%   \includegraphics[scale=0.28,angle=-90]{fig14}
  }
  \subfigure[$\alpha_1=16,\beta_1=16,\alpha_2=16,\beta_2=49$]{
%   \includegraphics[scale=0.28,angle=-90]{fig15}
  }
  \subfigure[$\alpha_1=1,\beta_1=16,\alpha_2=1,\beta_2=49$]{
%   \includegraphics[scale=0.28,angle=-90]{fig16}
  }
  \caption{Histograms constructed from samples with 10,000 elements, generated from the mixture of two Beta distributions with $w=0.5$. Below each histogram are the parameters of the mixture.}
  \label{fig:13141516}
\end{figure}

\begin{table}[h]
 \centering
 \caption{Efficiency comparison of SM and VLB methods for parameters covering uniformly the support of $B$. Column $n$ shows the number of parameter combinations with which each method has superior compression.}
 \begin{tabular}{ccc}
  \hline 
  Methods  & n   & Percentage \\
  \hline
  SM	   & 592	& 0.9034\% \\
  VLB	   & 64944	& 99.0966\% \\
  \hline
  Total    & 65536	& 100\% \\
  \hline
 \end{tabular}
 \label{tab:02}
\end{table}

\begin{figure}[h]
  \centering
  \subfigure[Efficiency histogram of the SM method ]{
%   \includegraphics[scale=0.25,angle=-90]{fig17}
  \label{fig:17}
  }
  \subfigure[Efficiency histogram of the VLB method ]{
%   \includegraphics[scale=0.25,angle=-90]{fig18}
  \label{fig:18}
  }
  \caption{Efficiency histograms of the SM and VLB methods. Note that the VLB method has a greater average efficiency than SM method .}
  \label{fig:1718}
\end{figure}

As we have shown, the VLB method is more effective compressing most integer datasets up to 64 bits in size. This is due to its ability to exploit the variance in the data set and reduce the waste of bits in the representation of some numbers. In specific cases where the variance in the data null or too small, method I will be more efficient. As a matter of fact, for matrices where all elements have the same bit-length, SM method  will always be better, regardless of bit-length (Figures \ref{fig19} and \ref{fig20}), The only exception if for bit-length 64 where neither method is able to compress the data.

\begin{figure}[h]
  \centering
%   \subfigure[Efficiencies of the SM and VLB methods, for constant bit-length matrices.]{\includegraphics[scale=0.3,clip]{fig19}\label{fig19}}
%   \subfigure[$\eta_1-\eta_2$ for matrices of constant bit-length.]{\includegraphics[scale=0.3,clip]{fig20}\label{fig20}}
  \caption{Compression efficiency of the SM and VLB methods for matrices of constant bit-length.}
  \label{fig:1920}
\end{figure}

@@

To determine the best compression method to apply, it's necessary to inspect the distribution of bit-lengths of matrix elements. When matrix elements are small or have nearly-constant bit-length, the SM Method is best, otherwise, the VLB method should be chosen.

As an example, let $M_{r\times c}$ be a integer matrix such that the half of its elements have bit-length $1$ and the other half $64$. Recalling Equation \ref{eq:17}, now we have two groups of elements (by bit-length), $b_1=1 $, $b_2=64$ and $f_i = \frac{rc}{2}$ for $i = 1 \text{ and } 2$. As the greatest bit-length is $64$, then $k=7$. Compression efficiency $\eta_2$ can be calculated using Equation \ref{eq:17}. After plugging in our numbers, we obtain a compression of $38.29\%$.

@@

The efficiency of the VLB method is influenced by the relative size of the bit-length groups. In this first example we considered only two groups, each comprised of half the matrix elements. Let's now vary the relative frequency of the groups, $\frac{f_i}{rc}$, while sticking to two groups. Let's also assume that $\frac{f_i}{rc}$ is a good approximation to the probability of a given bit-length in a matrix, which we will denote by $p_i$.

With this definition we can rewrite the Equation \ref{eq:17}, which becomes \ref{eq:29}. In Equation \ref{eq:29}, the $\frac{f_i}{rc}$ is replaced by $p_i$, representing the probability of elements from group $i$ in matrix M.

@@

Thus, when $p_1=0.1111$ and $p_2=0.8889$ the efficiency is minimal for the VLB method. For other combinations see Table \ref{tab:02}. Looking at this table, one can see two negative efficiencies, when ($p_1$,$p_2$) assume the values ​​(0,1) and (0.1,0.9). This correspond th cases when the method increases the memory requirements instead of decreasing it.

\begin{table}[h]
 \centering
 \caption{Combinations $p_1$ and $p_2$ to calculate the efficiency.}
 \begin{tabular}{ccc}
  \hline 
  $p_1$  & $p_2$ & $\eta_2$ \\
  \hline
  0.0	&1.0    &-0.109 \\
  0.1	&0.9	&-0.010 \\
  0.2	&0.8	&0.087 \\
  0.3	&0.7	&0.185 \\
  0.4	&0.6	&0.284 \\
  0.5	&0.5	&0.382 \\
  0.6	&0.4	&0.481 \\
  0.7	&0.3	&0.579 \\ 
  0.8	&0.2	&0.678 \\
  0.9	&0.1	&0.776 \\ 
  1.0	&0.0	&0.875 \\
  \hline
 \end{tabular}
 \label{tab:02}
\end{table}

So far, we have examined only two groups (hence two probabilities) of bit-length for the sake of simplicity. Before we generalize to probability distributions let's take a quick look at the efficiencies for more groups, with uniform probability:

@@

When the distribution of the group probabilities is uniform, i.e., the groups have approximately the same size, efficiency is basically the same, regardless of the number of groups.

Now we can leverage the notion of bit-length probabilities, and study efficiency when bit-lengths follow some  commonly used discrete probability distributions: Discrete Uniform, Binomial and Poisson. For all the experiments, we assume $k=7$, that is, the maximum possible bit-length is 64 bits. Thus, efficiency obtained will not be the best possible, since for that we would need assume small values of $k$ (Equation \ref{eq:29}).

\subsection*{Discrete Uniform}
Let the bit-lengths of the matrices be distribute according to the Uniform distribution $U(a=1,b=64)$, which means bit-lengths may take values in the set $\{1, 2, 3, $\ldots$, 64\}$ with equal probability, i.e., $\frac{1}{64}$.

@@

assuming all bit-lengths are possible, i.e., $a=1$ and $b=64$, and hence $k=7$, we can calculate $\eta_2$:

@@

This result agrees with the numerical estimates presented in Table \ref{tab:03}.

\paragraph{Numerical Estimates:}
To calculate the VLB efficiency, we generated a matrices with 100 ($M_{10 \times 10}$), 10,000 ($M_{100 \times 100}$) and 1,000,000 ($M_{1,000 \times 1,000}$) elements with 1, 8, 16, 32 and 64 number of bitss.  The average efficiency (Table \ref{tab:03}) is calculated from a 1,000 replicates of each matrix size. As expected the compression effiency gets better with lower expected bit-length.

\begin{table}[h]
  \centering
  \caption{Compression efficiency of VLB method of samples with bit-lengths coming from a Discrete Uniform distribution $U(a=1,b=64)$. Average efficiency ($\overline{\eta_2}\pm\textrm{SD}$) were calculated over a 1,000 replicates.}
 \begin{tabular}{cccc}
    \hline
    & &Matrix sizes& \\
    \hline
    & &Sample size & \\
    Expected bit-length & 100 & 10,000 & 1,000,000 \\
    \hline
     1&0.8750$\pm$0.0000& 0.8750$\pm$0.0000&0.8750$\pm$0.0000\\ 
     8&0.8202$\pm$0.0031& 0.8203$\pm$0.0003&0.8203$\pm$0.0000\\ 
     16&0.7580$\pm$0.0066& 0.7578$\pm$0.0007&0.7578$\pm$0.0001\\ 
     32&0.6330$\pm$0.0142& 0.6329$\pm$0.0014&0.6328$\pm$0.0001\\ 
     64&0.3826$\pm$0.0288& 0.3828$\pm$0.0028&0.3828$\pm$0.0003\\ 
    \hline
 \end{tabular}
 \label{tab:03}
\end{table}

@@

\paragraph{Theoretical Efficiency:}
Let bit-length $(B)$ be a random variable with Binomial distribution, $B \sim B(n=64,p=0.5)$, $E(b_i) = \sum_{i} b_i \times p(b_i) = n \times p$ and the eficiency becomes (with $k=7$):

@@

Which again agrees with estimates in Table \ref{tab:04}.

\paragraph{Numerical Estimates:}
For these experiments, the parameter n represents the maximum bit-length of matrix elements and takes values in $\{1, 8, 16, 32, 64\}$. In this case, we evaluate the efficiency as a function of the parameter $n$, and matrix size. Even though efficiency does not depend on matrix size, we tried different sizes to test the stability of the compression algorithm. Results are shown in Table \ref{tab:04}. As expected, smaller bit-lengths lead to higher compression efficiencies.

\begin{table}[h]
  \centering
  \caption{Compression efficiency with bit-lengths distributed according to a binomial distribution $B(n,0.5)$. Parameter $n \in \{1, 8, 16, 32, 64 \} $ represents the maximum bit-length. Since $p=0.5$ the expected bit-length is $n/2$ (first column)}
   \begin{tabular}{cccc}
     \hline
     & &Efficiency& \\
     \hline
     & &Sample size& \\
     Expected bit-length $(np)$ &100 &1,000 &1,000,000 \\
     \hline
     1 &0.8828$\pm$0.0004&0.8828$\pm$0.0004&0.8828$\pm$0.0000\\ 
     8 &0.8283$\pm$0.0022&0.8281$\pm$0.0002&0.8281$\pm$0.0000\\ 
     16&0.7656$\pm$0.0032&0.7656$\pm$0.0003&0.7656$\pm$0.0000\\ 
     32&0.6406$\pm$0.0045&0.6406$\pm$0.0004&0.6406$\pm$0.0000\\ 
     64&0.3910$\pm$0.0065&0.3906$\pm$0.0006&0.3906$\pm$0.0001\\
     \hline
  \end{tabular}
  \label{tab:04}
\end{table}

\subsection*{Poisson Distribution}
With bit-length derived from a Poisson($\lambda$), the parameter $\lambda$ corresponds to the expected bit-length. For the purpose of this analysis this Poisson distribution is truncated at 64. 
\paragraph{Theoretical Efficiency:}
Let bit-length $B \sim Poisson(\lambda=32)$. In this case, $E(b_i) = \lambda$, with $k=7$,  the efficiency becomes:

@@

This result is in accordance to Table \ref{tab:05}.
\paragraph{Numerical Estimates:}
The results for this simulation can be seen in \ref{tab:05}. Note that increasing the number of bits to represent numbers increases, there is a loss of efficiency in the compression process. In this case we did not simulate for $\lambda=64$ since a large portion of the samples would fall above the maximum bit-length we are considering for this analysis.

 \begin{table}[h]
   \centering
   \caption{Compression efficiency with bit-lengths distributed according to a Poisson distribution ($\lambda$), where $\lambda$ represents the expected bit-length.}
  \begin{tabular}{cccc}
      \hline
      &&Efficiency $\eta_2$      & \\
      \hline
      &&Matrix Size& \\
      Expected bit-length ($\lambda$)	& 100	& 1,000		    & 1,000,000 \\
      \hline
      1 	& 0.8751$\pm$0.0015 	& 0.8750$\pm$0.0004 & 0.8750$\pm$0.0000 \\ 
      8 	& 0.7654$\pm$0.0046 	& 0.7656$\pm$0.0005 & 0.7656$\pm$0.0000 \\ 
      16 	& 0.6405$\pm$0.0064 	& 0.6406$\pm$0.0006 & 0.6406$\pm$0.0001 \\ 
      32 	& 0.3908$\pm$0.0087 	& 0.3906$\pm$0.0009 & 0.3906$\pm$0.0001 \\ 
      \hline
  \end{tabular}
  \label{tab:05}
 \end{table}

@@

@@

\begin{figure}[htp]
  \includegraphics[scale=0.3,clip]{fig01.pdf}
  \caption{Representation of the compression process and allocation of 
elements of matrix M in the matrix m. The strip of value 700 is divided in two 
parts and stores in different strips in the matrix m.}
  \label{fig:01}
\end{figure}

@@

%\section*{Tables}
%\begin{table}[!ht]
%\caption{
%\bf{Table title}}
%\begin{tabular}{|c|c|c|}
%table information
%\end{tabular}
%\begin{flushleft}Table caption
%\end{flushleft}
%\label{tab:label}
% \end{table}

>>>>

 Where \textit{bits alocated} above mean total bits required for standard 
storage of the matrix, without compression, while bits used mean total bits 
requires to store the matrix after compression. From now on the efficiencies are 
denoted by $\eta_1$ for the SM method and by $\eta_2$ for the VLB method.

\subsubsection*{SM Method}

Let $M_{r \times c}$ be the matrix we wish to compress. In comparison with a 
conventional allocation (64-bit integers), we can apply Equation \ref{eq:14} to 
calculate the efficiency of the SM method:

@@

As we see in \ref{eq:15}, $\eta_1$ does not depend on size of the matrix, only 
on the bit-length of $\max M$. If $b(\max M)=64$, $\eta_1$ is 0, ie, no 
compression is possible. On the other extreme, if the matrix is composed 
exclusively of 0s and 1s, maximal compression is achievable, 
$\eta_1=0.9843$.\todo{Alterei a proporção para 0.98, pois é o maior valor que 
pode ser obtido.}

@@

Let the $rc$ elements of the matrix $M_{r \times c}$ be divided into $g$ groups, 
each with $f_i$ numbers of bit-length $b_i = b(m_i)$. Thus $f_i$ is the 
frequency of each bit-length present in $M$. Let $k = b(b(\max M))$, ie, the 
bit length of the bit-length of $\max M$. The efficiency $\eta_2$ is shown 
below.

@@

For this method, the highest value obtained for the data compression is 0.9687, 
when b = 1 and k = 1. However, the lower value is defined by a range of values, 
​​such that the value of $k = \frac{\sum_{i=1}^{g}  b_i \times f_i}{rc} - 64$. 
\todo{Adicionei esse trecho para ficar igual ao SM}

@@

\begin{equation}\label{eq:18}
 f(x) =  w \, Beta(\alpha_1,\beta_1) + (1-w)\,Beta(\alpha_2,\beta_2)
\end{equation}

The intention of using this mixture was to find a simple way to represent a 
large variety of bit length distributions. The first two central moments of this 
mixture are given in \ref{eq:19} and it will be used later to summarize our 
numerical results.

@@

From now on, we will apply Equations \ref{eq:15} and \ref{eq:17}, to determine 
the compression efficiency of SM and VLB methods for random matrices generated 
as describe above.

With $w=0$, a single Beta distribution is used. In Figure \ref{fig:02030405}, 
we show some distributions of bit-lengths for some combinations of $\alpha_1$ 
and $\beta_1$. From the figure it can be seen that a large variety of unimodal 
distributions can be generated in the interval $[1,64]$.

@@

Now let $w = 0.5$, ie, matrices will have elements with bit-lengths coming 
from a mixture of beta distributions, $B_1\sim Beta(\alpha_1,\beta_1)$ and 
$B_2\sim Beta(\alpha_2,\beta_2)$. The expected value for this mixture is shown 
in Equation \ref{eq:22}.

@@

As before, we generate 3 matrices of dimension $10,000$ for each 
parameterization, calculate the average efficiencies (Equations \ref{eq:23} and 
\ref{eq:24}) and their difference $D$.

Before moving on to efficiency results and analyses, let's first inspect samples 
from the mixture of transformed Beta distributions. Figures \ref{fig:10111213} 
and \ref{fig:14151617}, show a few parameterizations and their resulting sample 
distributions. It is important to note that from the mixture, we can now 
generate bimodal distributions as well as the unimodal types tested before. 
Since we are making statements about efficiency as a function of the expected 
bit-length, it is important to verify if these statements hold for bimodal  
distributions as well.

After sampling uniformly ($[1,5,9,\ldots,64]$, $n=65,536$) the bit-length space 
and comparing efficiencies, we summarized the results on Table \ref{tab:02}. In 
it we see how many parameterizations (from our sample) favor each method. We can 
also look at the distribution of efficiencies on our samples for each method 
(Figure \ref{fig:1819}), which clearly demonstrate the greater expected 
efficiency of method VLB (Figure \ref{fig:19}).

As we have shown, the VLB method is more effective compressing most integer 
datasets up to 64 bits in size. This is due to its ability to exploit the 
variance in the data set and reduce the waste of bits in the representation of 
some numbers. In specific cases, where the variance in the data null or too 
small, method I will be more efficient. As a matter of fact, for matrices where 
all elements have the same bit-length, SM method  will always be better, 
regardless of bit-length (Figures \ref{fig20} and \ref{fig21}). There are only 
exception, if for bit-length 64, where neither method is able to compress the 
data. \todo{Alterei a última frase}

@@

To determine the best compression method to apply, it's necessary to inspect the 
distribution of bit-lengths of matrix elements. When matrix elements are small 
or have nearly-constant bit-length, the SM Method is better\todo{alterei de 
best para better}, otherwise, the VLB method should be chosen.

As an example, let $M_{r\times c}$ be a integer matrix, such that the half of 
its elements have bit-length $1$ and the other half $64$. Recalling Equation 
\ref{eq:17}, now we have two groups of elements (by bit-length), $b_1=1 $, 
$b_2=64$ and $f_i = \frac{rc}{2}$ for $i = 1 \text{ and } 2$. As the greatest 
bit-length is $64$, then $k=7$. Compression efficiency $\eta_2$ can be 
calculated using Equation \ref{eq:17}. After plugging in our numbers, we obtain 
a compression of $38.29\%$.

@@

The efficiency of the VLB method is influenced by the relative size of the 
bit-length groups. In this first example, we considered only two groups, each 
comprised of half the matrix elements. Let's now vary the relative frequency of 
the groups, $\frac{f_i}{rc}$, while sticking to two groups. Let's 
also\todo{Acho que tinha que tirar esse Let's also, talvez colocar Now, we are 
going to} assume that $\frac{f_i}{rc}$ is a good approximation to the 
probability of a given bit-length in a matrix, which we will denote by $p_i$.

With this definition, we can rewrite the Equation \ref{eq:17}, which becomes 
\ref{eq:29}. In Equation \ref{eq:29}, the $\frac{f_i}{rc}$ is replaced by $p_i$, 
representing the probability of elements from group $i$ in matrix M.

@@

Thus, when $p_1=0.1111$ and $p_2=0.8889$ the efficiency is minimal for the VLB 
method. For other combinations see Table \ref{tab:03}. Looking at this table, 
one can see two negative efficiencies, when ($p_1$,$p_2$) assume the values 
​​(0,1) and (0.1,0.9). This correspond th cases, when the method increases, the 
memory requirements instead of decreasing it.

So far, we have examined only two groups (hence two probabilities) of bit-length 
for the sake of simplicity. Before, we generalize to probability distributions 
let's take a quick look at the efficiencies for more groups, with uniform 
probability:

@@

When the distribution of the group probabilities is uniform, ie, the groups 
have approximately the same size, efficiency is basically the same, regardless 
of the number of groups.

Now, we can leverage the notion of bit-length probabilities, and study 
efficiency when bit-lengths follow some commonly used discrete probability 
distributions: Discrete Uniform, Binomial and Poisson. For all the experiments, 
we assume $k=7$, that is, the maximum possible bit-length is 64 bits. Thus, 
efficiency obtained will not be the best possible, since for that we would need 
assume small values of $k$ (Equation \ref{eq:29}).

\subsection*{Discrete Uniform}
Let the bit-lengths of the matrices be distribute according to the Uniform 
distribution $U(a=1,b=64)$, which means bit-lengths may take values in the set 
$\{1, 2, 3, $\ldots$, 64\}$ with equal probability, ie, $\frac{1}{64}$.

@@

\noindent assuming all bit-lengths are possible, i.e., $a=1$ and $b=64$, and 
hence $k=7$, we can calculate $\eta_2$:

@@

This result agrees with the numerical estimates presented in Table \ref{tab:04}.

\paragraph{Numerical Estimates:}

To calculate the VLB efficiency, we generated a matrices with 100 ($M_{10 \times 
10}$), 10,000 ($M_{100 \times 100}$) and 1,000,000 ($M_{1,000 \times 1,000}$) 
elements with 1, 8, 16, 32 and 64 number of bits.  The average efficiency (Table 
\ref{tab:03}) is calculated from a 1,000 replicates of each matrix size. As 
expected the compression efficiency gets better with lower expected bit-length.

@@

\paragraph{Theoretical Efficiency:}
Let bit-length $(B)$ be a random variable with Binomial distribution, $B \sim 
B(n=64,p=0.5)$, $E(b_i) = \sum_{i} b_i \times p(b_i) = n \times p$ and the 
efficiency becomes (with $k=7$):

@@

Which again agrees with estimates in Table \ref{tab:05}.

\paragraph{Numerical Estimates:}

For these experiments, the parameter n represents the maximum bit-length of 
matrix elements and takes values in $\{1, 8, 16, 32, 64\}$. In this case, we 
evaluate the efficiency as a function of the parameter $n$, and matrix size. 
Even though efficiency does not depend on matrix size, we tried different sizes 
to test the stability of the compression algorithm. Results are shown in Table 
\ref{tab:05}. As expected, smaller bit-lengths lead to higher compression 
efficiencies.

\subsection*{Poisson Distribution}
With bit-length derived from a Poisson($\lambda$), the parameter $\lambda$ 
corresponds to the expected bit-length. For the purpose of this analysis this 
Poisson distribution is truncated at 64.

\paragraph{Theoretical Efficiency:}
Let bit-length $B \sim Poisson(\lambda=32)$. In this case, $E(b_i) = \lambda$, 
with $k=7$,  the efficiency becomes:

@@

This result is in accordance to Table \ref{tab:06}.

\paragraph{Numerical Estimates:}
The results for this simulation can be seen in \ref{tab:06}. Note that 
increasing the number of bits to represent numbers increases, there is a loss of 
efficiency in the compression process. In this case, we did not simulate for 
$\lambda=64$ since a large portion of the samples would fall above the maximum 
bit-length, but \todo{coloquei ", but", aqui} we are considering for this 
analysis.

@@

\todo{Parei aqui}

@@

\begin{figure}[h]
  \includegraphics[scale=0.3,clip]{fig01}
  \caption{Representation of the compression process and allocation of 
elements of matrix M in the matrix m. The strip of value 700 is divided in two 
parts and stores in different strips in the matrix m.}
  \label{fig:01}
\end{figure}

\begin{figure}[h]
  \centering
  \subfigure[$\alpha=1,\beta_1=1$]{
   \includegraphics[scale=0.25,angle=-90,clip]{fig02}
  }
  \subfigure[$\alpha=1,\beta=32$]{
   \includegraphics[scale=0.25,angle=-90,clip]{fig03}
  }
  \subfigure[$\alpha=32,\beta=1$]{
   \includegraphics[scale=0.25,angle=-90,clip]{fig04}
  }
  \subfigure[$\alpha=64,\beta=64$]{
   \includegraphics[scale=0.25,angle=-90]{fig05}
  }
  \caption{Histograms constructed from samples with 10,000 elements, generated 
from a Beta distribution. Below each histogram is possible to verify the 
parameters used.With these two parameters it is possible to perform different 
combinations of numbers to fill the arrays to be compressed.}
  \label{fig:02030405}
\end{figure}\todo{Adicionei a última frase}

\begin{figure}[h]
  \centering
   \subfigure[SM-VLB  Difference ($D = \eta_1 - \eta_2$)]{ 
    \includegraphics[scale=0.19,clip]{fig08}
   }
   \subfigure[SM Efficiency  ($\eta_1$)]{
    \includegraphics[scale=0.19,clip]{fig07}
   }
   \subfigure[VBL Efficiency ($\eta_2$)]{
    \includegraphics[scale=0.19,clip]{fig06}
   }
   \subfigure[SM$\geq$VLB Efficiency ($\eta_1 \geq \eta_2$)]{
    \includegraphics[scale=0.19,clip]{fig09}
   }
  \caption{Comparing compression efficiency of methods 1 and 2. Color scale in 
represent average bit-length. In (a) we can see the difference $D = \eta_1 - 
\eta_2$. It can be seen that for most combination of $\alpha$ and $\beta$, 
$D<0$, meaning the second method is more efficient to compress a sample of 
numbers with bit-lengths coming from a $Beta(\alpha,\beta)$ distribution. 
However, there is a small region in parameter space, which is shown in white on 
(d), where the SM method is more efficient. This region corresponds to the dots 
in red in (a), where the average bit-length is higher. In panels (b) and (c), we 
can see the efficiencies of SM and VLB methods, respectively.}
  \label{fig:06070809}
\end{figure}
\todo{A escala de cor é a mesma em todas as figuras, no texto estava dizendo 
que era apenas na a, b e c. Eu alterei.}

\begin{figure}[h]
  \centering
  \subfigure[$\alpha_1=1,\beta_1=1,\alpha_2=1,\beta_2=1$]{
   \includegraphics[scale=0.28,angle=-90,clip]{fig10}
  }
  \subfigure[$\alpha_1=1,\beta_1=32,\alpha_2=32,\beta_2=1$]{
   \includegraphics[scale=0.28,angle=-90,clip]{fig11}
  }
  \subfigure[$\alpha_1=32,\beta_1=32,\alpha_2=32,\beta_2=32$]{
   \includegraphics[scale=0.28,angle=-90,clip]{fig12}
  }
  \subfigure[$\alpha_1=64,\beta_1=32,\alpha_2=32,\beta_2=64$]{
   \includegraphics[scale=0.28,angle=-90]{fig13}
  }
  \caption{Histograms constructed from samples with 10,000 elements, generated 
from the mixture of two Beta distributions with $w=0.5$. Below each histogram 
are the parameters of the mixture.}
  \label{fig:10111213}
\end{figure}

\begin{figure}[h]
  \centering
  \subfigure[$\alpha_1=64,\beta_1=48,\alpha_2=1,\beta_2=48$]{
   \includegraphics[scale=0.28,angle=-90]{fig14}
  }
  \subfigure[$\alpha_1=16,\beta_1=46,\alpha_2=49,\beta_2=64$]{
   \includegraphics[scale=0.28,angle=-90]{fig15}
  }
  \subfigure[$\alpha_1=16,\beta_1=16,\alpha_2=16,\beta_2=49$]{
   \includegraphics[scale=0.28,angle=-90]{fig16}
  }
  \subfigure[$\alpha_1=1,\beta_1=16,\alpha_2=1,\beta_2=49$]{
   \includegraphics[scale=0.28,angle=-90]{fig17}
  }
  \caption{Histograms constructed from samples with 10,000 elements, generated 
from the mixture of two Beta distributions with $w=0.5$. Below each histogram 
are the parameters of the mixture.}
  \label{fig:14151617}
\end{figure}

\begin{figure}[h]
  \centering
  \subfigure[Efficiency histogram of the SM method ]{
   \includegraphics[scale=0.25,angle=-90]{fig18}
  \label{fig:18}
  }
  \subfigure[Efficiency histogram of the VLB method ]{
   \includegraphics[scale=0.25,angle=-90]{fig19}
  \label{fig:19}
  }
  \caption{Efficiency histograms of the SM and VLB methods. Note that the VLB 
method has a greater average efficiency than SM method.}
  \label{fig:1819}
\end{figure}

\begin{figure}[h]
  \centering
   \subfigure[Efficiencies of the SM and VLB methods, for constant bit-length 
matrices.]{\includegraphics[scale=0.3,clip]{fig20}\label{fig20}}
   \subfigure[$\eta_1-\eta_2$ for matrices of constant 
bit-length.]{\includegraphics[scale=0.3,clip]{fig21}\label{fig21}}
  \caption{Compression efficiency of the SM and VLB methods for matrices of 
constant bit-length.}
  \label{fig:2021}
\end{figure}

@@

\section*{Tables}
%\begin{table}[!ht]
%\caption{
%\bf{Table title}}
%\begin{tabular}{|c|c|c|}
%table information
%\end{tabular}
%\begin{flushleft}Table caption
%\end{flushleft}
%\label{tab:label}
% \end{table}

\begin{table}[h]
 \centering
 \caption{Efficiency comparison of SM and VLB methods for parameters covering 
uniformly the support of $B$. Column $n$ shows the number of parameter 
combinations with which each method has superior compression.}
 \begin{tabular}{crr}
  \hline 
  Methods  & n   & Percentage \\
  \hline
  SM	   & 592	& 0.9034\% \\
  VLB	   & 64944	& 99.0966\% \\
  \hline
  Total    & 65536	& 100\% \\
  \hline
 \end{tabular}
 \label{tab:02}
\end{table}

\begin{table}[h]
 \centering
 \caption{Combinations $p_1$ and $p_2$ to calculate the efficiency.}
 \begin{tabular}{ccc}
  \hline 
  $p_1$  & $p_2$ & $\eta_2$ \\
  \hline
  0.0	&1.0    &-0.109 \\
  0.1	&0.9	&-0.010 \\
  0.2	&0.8	&0.087 \\
  0.3	&0.7	&0.185 \\
  0.4	&0.6	&0.284 \\
  0.5	&0.5	&0.382 \\
  0.6	&0.4	&0.481 \\
  0.7	&0.3	&0.579 \\ 
  0.8	&0.2	&0.678 \\
  0.9	&0.1	&0.776 \\ 
  1.0	&0.0	&0.875 \\
  \hline
 \end{tabular}
 \label{tab:03}
\end{table}
